Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-13 manuscript/manuscript.tex
Update grid search text

We compute the relative difference between the numerical and analytical results for the
gravitational potential ($V$) and the vertical component of its gradient ($g_z$) on the
computation grids detailed in Table~\ref{tab:grids} and using shell models with the same
thicknesses as the linear density case.
Because this is a time consuming computation, we limit the analysis only to the second
collection of tesseroids in Table~\ref{tab:tess-collections}.
For the sake of brevity, we compute the maximum differences between the results
involving each computation grid, shell thickness and value of the $b$ factor.
The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
lines are the ones that present a numerical error lower than the 0.1\% threshold.
Also shown in Fig.~\ref{fig:grid-search} are the $D$ values determined for the linear
density function in the previous section ($D_\text{linear}$).

The minimum values of $D$ that are within the 0.1\% threshold are the $D_\text{linear}$
ones, i.e.~they are the optimal values of the distance-size ratios for $V$ and $g_z$.
These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
to the exponential case.
This conclusion can be understood considering that the two dimensional adaptive
discretization and the density-based algorithms are independent of each other: the first
one splits the tesseroid only in the latitudinal and longitudinal directions, while the
latter only does it in the radial direction.
This leads to the fact that the value of $\delta$ needed to limit the numerical error
due to the density variation is independent of the value of $D$ and vice versa.
That's why the optimal distance-size ratios for the exponential density case is the same
as the ones for the linear density case.
Because the grid search was limited to specific values of $b$ and used only one grid, we
will determine the optimal value of $\delta$ for exponential density on the next
subsection.

\begin{figure}
\centering
\iftwocol{
\includegraphics[width=0.9\linewidth]
    {figures/grid-search.pdf}
}{
\includegraphics[width=0.5\linewidth]
    {figures/grid-search.pdf}
}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison between the
    analytical solution and the numerical approximation of the gravitational fields ($V$
    and $g_z$) generated by a spherical shell with an exponential density function
    (Eq.~\ref{eq:density-exp}).
    These comparisons were carried out on the ``Global'' grid (Table~\ref{tab:grids}),
    with only five spherical shell models made out of the second tesseroid collection
    from Table~\ref{tab:tess-colections} with the different thicknesses defined in the
    same Table, and density function with $b = 30$.
    The percentage difference values are obtained as the maximum difference
    corresponding to each spherical shell model.
    The points inside the dashed line are the ones that present an error lower than
    0.1\%.
    }
\label{fig:grid-search}
\end{figure}

>>>>

We compute the relative difference between the numerical and analytical results for the
gravitational potential ($V$) and the vertical component of its gradient ($g_z$) on the
computation grids detailed in Table~\ref{tab:grids}.
Because this is a time consuming computation, we limit the analysis only to five
spherical shell models made out of the second tesseroid collection from
Table~\ref{tab:tess-collections} with the five different thicknesses, and only one value
of $b$ equal to 30 (Fig.~\ref{fig:exp-densities}).
For the sake of brevity, we compute the maximum differences between the results
involving each shell thickness.
The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
lines are the ones that present a numerical error lower than the 0.1\% threshold.
Also shown in Fig.~\ref{fig:grid-search} are the $D$ values determined for the linear
density function in the previous section ($D_\text{linear}$).

The minimum values of $D$ that are within the 0.1\% threshold are the $D_\text{linear}$
ones, i.e.~they are the optimal values of the distance-size ratios for $V$ and $g_z$.
These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
to the exponential case.
This conclusion can be understood considering that the two dimensional adaptive
discretization and the density-based algorithms are independent of each other: the first
one splits the tesseroid only in the latitudinal and longitudinal directions, while the
latter only does it in the radial direction.
This leads to the fact that the value of $\delta$ needed to limit the numerical error
due to the density variation is independent of the value of $D$ and vice versa.
That's why the optimal distance-size ratios for the exponential density case is the same
as the ones for the linear density case.

Because the grid search was limited to specific values of $b$ and used only one grid, we
will determine the optimal value of $\delta$ for exponential density tesseroids on the
next subsection.

\begin{figure}
\centering
\iftwocol{
\includegraphics[width=\linewidth]
    {figures/grid-search.pdf}
}{
\includegraphics[width=0.5\linewidth]
    {figures/grid-search.pdf}
}
\caption{
    Numerical error exploration in the $D$-$\delta$ space.
    The percentage difference values were obtained from the comparison between the
    analytical solution and the numerical approximation of the gravitational fields ($V$
    and $g_z$) generated by a spherical shell with an exponential density function
    (Eq.~\ref{eq:density-exp}).
    These comparisons were carried out on the ``Global'' grid (Table~\ref{tab:grids}),
    with only five spherical shell models made out of the second tesseroid collection
    from Table~\ref{tab:tess-colections} with the different thicknesses defined in the
    same Table, and density function with $b = 30$.
    The percentage difference values are obtained as the maximum difference
    corresponding to each spherical shell model.
    The points inside the dashed line are the ones that present an error lower than
    0.1\%.
    }
\label{fig:grid-search}
\end{figure}
